---
output: github_document
bibliography: paper.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# implicitMeasures

R package for computing different Implicit Measures scores
  <!-- badges: start -->
[![Travis build status](https://travis-ci.org/OttaviaE/implicitMeasures.svg?branch=master)](https://travis-ci.org/OttaviaE/implicitMeasures)
  <!-- badges: end -->
  
## Aim and Overview

The goal of `implicitMeasures` package is to provide a tool for easily cleaning and computing the scores for the Implicit Association Test [IAT; @Greenwald:1998] and the Single Catgeory Implicit Association Test [SC-IAT; @karpinski2006]. 

Six different algortihms for the computation of the IAT effect, the so-called *D-score*, are available [@Greenwald2003]. The six different alogrithms differentiate themselves for the treatment of the extreme fast responses and for the strategy used for replacing the error responses. Different mistakes can be made during the computation of the *D-score*, which can raise replicability issues of the results. On top of that, many reserahcres fail to report the exact algorithm they have used for computing the *D-score*, raising again replicability issues [@ellithorpe2015]. 

`implicitMeasures` include the following functions: 

* `clean_iat()`: Prepare and clean IAT data.
* `clean_sciat()`: Prepare and clean SC-IAT data.
* `computeD()`: Compute IAT *D-score*.
* `Dsciat()`: Compute SC-IAT *D-score*.
* `descript_d()`: Descriptive table of *D-score*s (also in  LaTeX) format.
*`d_distr()`: Plot IAT or SC-IAT scores (distribution).
* `d_plot()`: Plot IAT or SC-IAT scores (points).
* `multi_dscore()`: Compute and plot multiple IAT *D-score*s.
* `multi_dsciat()`: Plot SC-IAT *D-score*s.
* `IATrel()` computes the IAT reliablity [@gaw2017]

All the functions for graphically represernting the results are based on `ggplot2` [@Wickham2016].

## Installation

You can install the released version of `implicitMeasures` from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("implicitMeasures")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("OttaviaE/implicitMeasures")
```
## Example

This is a basic example which shows you how to compute the IAT *D-score*. More detailed examples are given the package vignettes.

```{r example}
library(implicitMeasures)
## load the raw_data dataframe
data("raw_data")

## prepare the dataset for the computation
iat_data <- clean_iat(raw_data, 
                          sbj_id = "Participant",
                          block_id = "blockcode",
                          mapA_practice = "practice.iat.Milkbad",
                          mapA_test = "test.iat.Milkbad",
                          mapB_practice = "practice.iat.Milkgood",
                          mapB_test = "test.iat.Milkgood",
                          latency_id = "latency",
                          accuracy_id = "correct",
                          trial_id = "trialcode",
                          trial_eliminate = c("reminder", "reminder1"))

## store the dataset for comptuing the D-score
iat <- iat_data[[1]]

## Compute the D-score
dscore <- computeD(iat, D = "d3")
```

`computeD()` results in a data frame with class `dscore`, and it can be passed to other functions, for example for plotting the results, either at the individual level: 

```{r pointsExample, echo=FALSE, fig.align='center', message=FALSE, warning=FALSE}
d_plot(dscore, 
       x_values = FALSE)
```
or at the sample level: 

```{r densityexamples, echo =FALSE, fig.align='center', message=FALSE, warning=FALSE}
d_distr(dscore, 
        graph = "density")
```

## Data import

You can import import your data in any format you want. If you are importing data sets from SPSS, please use either `haven::read_sav("~/path/to/mydata.sav")` or `foreign::read.spss("~/path/to/mydata.sav")` without changing the default options of the functions. `implicitMeasures` will recognize that the data frame is coming from SPSS and will handle that. 

## Bugs and problems

If you find any bugs or encounter any problems in using this package, please post a minimal reproducible example on [github](https://github.com/OttaviaE/implicitMeasures/issues). 
For questions and other discussion, you can contact the author and maintainer of the package at otta.epifania@gmail.com.

# References
