---
title: "IAT-example"
author: Ottavia M. Epifania
date: "`r Sys.Date()`"
bibliography: vignette.bib
output:  rmarkdown::html_vignette
#  pdf_document: default
vignette: >
  %\VignetteIndexEntry{IAT-example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", 
  #fig.path = "",
  warning = FALSE, 
  message = FALSE
)
```

```{r setup}
library(implicitMeasures)
```

This vignette illustrates how to use `implicitMeasures` package for computing IAT *D-score*. The illustration is based on the data set `raw_data` that is provided with the package.

### First thing first: Import and explore data
Labels containing specification `.iat` in variable `blockcode` identify IAT blocks.

```{r}
data("raw_data")
# explore the dataframe
str(raw_data)
# explore the levels of the blockcode variable to identify the IAT blocks
levels(raw_data$blockcode)
```

Once the IAT blocks have been identified, it is possible to prepare and clean the IAT data by using function `clean_iat()`. Since the data set also includes respondents' demographic information (`demo` label in `blockcode` variable), it is possible to extract and store these information in a separate data frame:  

```{r eval = T}
iat_cleandata <- clean_iat(raw_data, sbj_id = "Participant",
                          block_id = "blockcode",
                          mapA_practice = "practice.iat.Milkbad",
                          mapA_test = "test.iat.Milkbad",
                          mapB_practice = "practice.iat.Milkgood",
                          mapB_test = "test.iat.Milkgood",
                          latency_id = "latency",
                          accuracy_id = "correct",
                          trial_id = "trialcode",
                          trial_eliminate = c("reminder", "reminder1"),
                          demo_id = "blockcode",
                          trial_demo = "demo")
```

Since also the demographic data has been specified, `clean_iat()` results in a list of 3 elements: 

```{r}
str(iat_cleandata)
```

`data_keep` is a `data.frame` with class `iat_clean`. It contains the data set to pass to function `compute_iat()`. 

`data_eliminate` is a `data.frame` containing all discarded blocks and trials. `demo` is a `data.frame` containing all the trials identified as `demo` in the `blockcode` variable.

Store the first `data_keep` element in a data frame to pass to function `compute_iat()`. 

```{r}
iat_data <- iat_cleandata[[1]]
head(iat_data)
```

You can also skip this step and directly pass `iat_cleandata[[1]]` to function `compute_iat()`.

### Compute the IAT D-score

Once that IAT data have been prepared and cleaned by function `clean_iat()`, it is possible to compute the *D-score* by using function `compute_iat()`.

This function only takes two arguments. The first argument is the data frame with class `iat_clean`, the second argument is a character specifying the *D-score* algorithm for the computation.  To compute multiple *D-score* algorithms at the same time, use `multi_dscore()` function. 

```{r}
dscore <- compute_iat(iat_data, Dscore = "d3")
str(dscore)
```

Function `compute_iat()` results in a `data.frame` with class `dscore` containing a number of rows equal to the number of participants, their *D-score*s and a bunch of useful information on their performance (see the documentation of the `compute_iat()` function for further details on the resulting data frame). Functions `IAT_rel()`, `descript_d()`, `d_point()`, and `d_density()` require the object resulting from function `compute_iat()`. 

### IAT descriptive statistics and reliability

The descriptive statistics of the *D-score*, computed on the practice and test blocks, and the actual *D-score*, can be easily obtained with function `descript_d()`: 

```{r}
descript_d(dscore) # Data frame containing IAT D-scores
```

By specifying `latex = TRUE`, function `descript_d()` print the results in LaTeX code: 

```{r}
descript_d(dscore, # Data frame containing IAT D-scores
           latex = TRUE) # obtain the code for latex tables
```

Function `IAT_rel()` computes IAT reliability by correlating the *D-score* obtained from practice blocks with the *D-score* obtained from test blocks [see @gaw2017 for further details]: 

```{r}
IAT_rel(dscore)
```

### Plotting the results

`implicitMeasures` comes with several functions for obtaining nice and clear representations of the results at both individual respondent and sample levels. Additionally, it includes a function for computing and plotting multiple IAT *D-score* algorithms at the same time. 

#### Individual respondent level

Function `d_point()` plots the IAT *D-score* for each respondent. 

```{r, fig.align='center', fig.width=6, fig.cap="d_point() function with default settings"}
d_point(dscore) # Data frame containing IAT D-scores
```
Clearly, in case of large sample size, the label identifying each respondent is not easy to read, and it can be eliminated by setting argument `x_values = FALSE`. 
Respondents can be arranged by increasing or decreasing *D-score* by setting argument `order_sbj` equal to either `"D-increasing"`  or `"D-decreasing"`. Descriptive statistics ($M_{\text{D-score}}\pm 2sd$) can be added by setting the argument `include_stats = TRUE`. Finally, the color of the points can be changed by using argument `col_point`. 

```{r, fig.align='center', fig.width=6, fig.cap="d_point() function with   settings change"}
d_point(dscore, # dataframe containing IAT D-scores
       order_sbj = "D-decreasing", # change respondents order
       x_values = FALSE, # remove respodents' labels
       include_stats = TRUE, # include descriptive statistics
       col_point = "lightskyblue") # change points color
```



#### Sample level 

Function `d_density()` plots the distribution of the IAT *D-score*s. It provides different options for choosing the most appropriate representation. 


```{r, fig.align='center', fig.width=6, fig.cap="d_density() function with default settings"}
d_density(dscore) # dataframe containing IAT Dscores
```

The number of bins can be changed with argument `n_bin`. Argument `graph` can be used for changing the graphical representation of the data. 
It is possible to choose an histogram representation (`graph = "histogram"`, default), a representation of the density distribution (`graph = "density"`), or a violin plot (`graph = "violin"`). Argument `col_fill` can be used to change the color of the points representing each respondent's score in the violin plot. 
Finally, also descriptive statistics (i.e., $M_{\text{D-score}}  \pm 2sd$) can be added to the graph by setting argument `include_stats = TRUE`. 

```{r sampleSettings, fig.align='center', fig.width=6, fig.cap="\\label{fig:sampleSettings}d_density() function with settings change"}
d_density(dscore, # dataframe containing IAT Dscores
        graph = "violin", # change graphical representation
        include_stats = TRUE) # include descriptive statistics
```


#### Multiple D-scores

Function `multi_dscore()` computes and plot multiple *D-score* algorithms. The *D-score* algorithms that can be computed depend on IAT administration. If the IAT administration included a feedback strategy (i.e., built-in correction), only *D1* and *D2* algorithms should be computed. If the IAT administration did not include a feedback strategy, then algorithms *D3*, *D4*, *D5*, and *D6* should be computed. An exhaustive and detailed illustration of the *D-score* algorithms is provided in the "Implicit Measures" vignette. To specify the algorithms that can be computed, argument `ds` must be set equal to either `"built-in"` (for the computation of *D1* and *D2*) or `error-inflation` (for the computation of all other algorithms). 

```{r}
multi_scores <- multi_dscore(iat_data, # object with class "iat_clean"
                             ds = "error-inflation") # string specifying the 
                                            # algorithms to compute
```

Function `multi_dscore()` results in a `list` containing two objects. The first object is a `data.frame` containing all computed algorithms and respondents' IDs. 

```{r}
multi_d <- multi_scores[[1]]
head(multi_d)
str(multi_d)
```

The second object is a `ggplot` graph displaying the distribution of the *D-score*s computed with each algorithm in violin plots.

```{r, fig.align='center', fig.width=6, fig.cap="Multiple D-scores representation"}
multi_scores[[2]]
```

# References
