---
title: "IAT-example"
author: Ottavia M. Epifania
date: "`r Sys.Date()`"
bibliography: vignette.bib
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{IAT-example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(implicitMeasures)
```

This vignette illustrates how to use `implicitMeasures` package to compute the IAT *D-score* by using the `raw_data` data set. 

### First thing first: Import and explore data
Labels containing the `.iat` specification in the `blockcode` variable identity the IAT blocks.

```{r}
data("raw_data")
# explore the dataframe
str(raw_data)
# explore the levels of the blockcode variable to identify the IAT blocks
levels(raw_data$blockcode)
```

Once the IAT blocks have been identified, it is possible to prepare and clean the IAT data by means of the `clean_iat()` function. Since the data set also includes respondents' demographic information (`demo` label in `blockcode` variable), it is possible to extract and store these information in a separate data frame:  

```{r eval = T}
iat_cleandata <- clean_iat(raw_data, sbj_id = "Participant",
                          block_id = "blockcode",
                          mapA_practice = "practice.iat.Milkbad",
                          mapA_test = "test.iat.Milkbad",
                          mapB_practice = "practice.iat.Milkgood",
                          mapB_test = "test.iat.Milkgood",
                          latency_id = "latency",
                          accuracy_id = "correct",
                          trial_id = "trialcode",
                          trial_eliminate = c("reminder", "reminder1"),
                          demo_id = "blockcode",
                          trial_demo = "demo")
```

Since the also the demographic data has been specified, `clean_iat()` results in a list of 3 elements: 

```{r}
str(iat_cleandata)
```

`data_keep` is a `data.frame` with class `iat_clean`. It contains the data to pass to the `computeD()` function. 

`data_eliminate` is a `data.frame`containing all the discarded blocks and trials, while `demo`is a `data.frame` containing all the trials identified as `demo` in the `blockcode` variable.

Store the first `data_keep` in a data frame to pass to the `computeD()` function. 

```{r}
iat_data <- iat_cleandata[[1]]
head(iat_data)
```
### Compute the D-score

Once the IAT have been prepared and cleaned through the `clean_iat` function, it is possible to compute the *D-score* by using the `computeD()` function.

This function only takes two elements: The first element is a data frame with class `iat_clean`, the second element is the *D-score* one wants to compute. For computing multiple *D-score* algorithms at the same time, use the `multi_dgraph()` function. 

```{r}
dscore <- computeD(iat_data, Dscore = "d3")
str(dscore)
```

The `computeD()` function results in a `data.frame` with class `dscore` containing a number of rows equal to the number of participants, their *D-score*s and a bunch of useful information on their performance (see the documentation of the `computeD()` function for further details on the resulting data frame). Functions `IATrel()`, `dscr_d()`, `d_plot()`, and `d_distr()` require the object resulting from the `computeD` function. 

### IAT descriptive statistics and reliability

The descriptive statistics of the *D-score*, computed on the practice and test blocks and the actual *D-score*, can be easily obtained by means of the `dscr_d()` function: 

```{r}
descript_d(dscore)
```

By specifying `latex = TRUE`, the `dscr_d()` print the results in Latex: 

```{r}
descript_d(dscore, latex = T)
```

The `IATrel()` computes the IAT reliability as the correlation between the *D-score*s computed on the practice and test blocks across participants [see @gaw2017 for further details]: 

```{r}
IATrel(dscore)
```

### Plotting the results

`implicitMeasures` provides the users with different functions for plotting the results, either at the individual or at the sample level. Furthermore, it gives the chance to compute and plot multiple *D-score* algorithms at the same time. 

#### Individual level

`d_plot()` plots the *D-score* for each participant in the sample. Participants can follow the order they had in the original data frame (`default`), or they can be ordered by increasing or decreasing *D-score* (`D-increasing` and `D-decreasing`, respectively). The values labels of the x-axis can be removed by setting `x_values = FALSE` (suggested in case of big sample size). The `include_stats` (default = `FALSE`) argument plots the descriptive statistics of the *D* (mean +/- 2sd). 

```{r, fig.align='center', fig.width=6}
d_plot(dscore)
# change respondents order, remove x-values
d_plot(dscore, order_sbj = "D-increasing", 
       x_values = FALSE)
# change respondents order, remove x-values, and add decsriptive statistics
d_plot(dscore, order_sbj = "D-decreasing", 
       x_values = FALSE, include_stats = TRUE)
```

Points color can be changed as well by setting `col_point` argument equal to the desired color (e.g., `col_point = "lightskyblue"`). 

#### Sample level 

`d_distr()` plots the distribution of the *D-score*s. It provides different options for choosing the most appropriate distribution by means of the `graph` argument, that can be set equal to `histogram` (histogram plot, default), `density` (density plot), or `violin` (violin plot). Descriptive statistics can be added as well (`include_stats = TRUE`). The filling color for the histogram and the density plot can be modified by means of the `col_fill` argument (e.g., `col_fill = "seagreen"`). The `col_point` argument can be used to change the color of the points in the violin plot (e.g., `col_fill = "royalblue"`).

```{r, fig.align='center', fig.width=6}
d_distr(dscore)
# change the number of bins
d_distr(dscore, n_bin = 120)
# change graph and add descriptive statistics
d_distr(dscore, graph = "density", include_stats = TRUE)
# change graph and add descriptive statistics
d_distr(dscore, graph = "violin", include_stats = TRUE)
```

#### Multiple D-scores

`multi_dscore()` computes and plot multiple *D-score* algorithms. The multiple *D-score*s that can be computed depend on IAT administration. If it included a feedback strategy, only *D1* and *D2* should be computed (`ds = "built-in"`), otherwise *D-score* from *D3* to *D6* should be computed (`ds = "error-inflation"`). 

```{r}
multi_scores <- multi_dscore(iat_data, ds = "error-inflation")
```

`multi_dscore()` results in a `list` containing two objects. The first object is a `data.frame` containing all the computed algorithms. 

```{r}
multi_d <- multi_scores[[1]]
head(multi_d)
str(multi_d)
```

The second object is a `ggplot` graph displaying the distribution of the algorithms.
```{r fig.align='center', fig.width=6}
multi_scores[[2]]
```

