---
title: "SC-IAT-example"
author: Ottavia M. Epifania
date: "`r Sys.Date()`"
bibliography: vignette.bib
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SC-IAT-example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(implicitMeasures)
```

This vignette illustrates how to use `implicitMeasures` package to compute the SC-IAT *D* by using the `raw_data` data set. 

### First thing first: Import and explore data

Labels containing the `sc_` specification in the `blockcode` variable identity the SC-IAT blocks.

```{r}
data("raw_data")
# explore the dataframe
str(raw_data)

# explore the levels of the blockcode variable to identify the SC-IAT blocks
levels(raw_data$blockcode)
```

`raw_data` contains data from two different SC-IATs, one for the assessment of the Milk chocolate preference (identified by the label `sc_milk`), and one for the assessment of the Dark chocolate preference (identified by the label `sc_dark`). 

Once the SC-IATs blocks have been identified, it is possible to prepare and clean the data for computing the *D*. Since the data set contains data from two SC-IATs, it is possible to separate them in two different data sets. Demographic information (label `demo` in `blockcode` variable) are available as well, and they can be extracted and stored in a separate data frame. It is not mandatory to have two SC-IATs or the demographic information, the function can work with just the first (and only) SC-IAT blocks specified.

```{r}
data("raw_data")
sciat_data <- clean_sciat(raw_data, sbj_id = "Participant",
                         block_id = "blockcode",
                         latency_id = "latency",
                         accuracy_id = "correct",
                         block_sciat_1 = c("test.sc_dark.Darkbad",
                                           "test.sc_dark.Darkgood"),
                         block_sciat_2 = c("test.sc_milk.Milkbad",
                                           "test.sc_milk.Milkgood"),
                         trial_id  = "trialcode",
                         trial_eliminate = c("reminder",
                                             "reminder1"), 
                         demo_id = "blockcode", 
                         trial_demo = "demo")
```

**DON'T USE THE `trial_eliminate` ARGUMENT TO ELIMINATE TRIALS EXCEEDDING THE RESPONSE TIME WINDOW (rtw).**

The labels for identifying the responses beyond the rtw (and that hence have to be eliminated) must be included in the varibale identified by `trial_id`, but they have to be specified via the `non_response` in the `Dsciat()` function to actually be deleted.

Since two SC-IATs and demographic data were specified, `clean_sciat()` results in a `List` of 3 elements: 

```{r}
str(sciat_data) # structure of the resulting List
```

The first two elements (`sciat1` and `sciat2`) are two  `data.frame` with class `sciat_clean` and they contain the data of the SC-IATs specified in the `block_sciat1` and `block_sciat2` arguments of the `clean_sciat()` function, respectively. The third element (`demo`) is a `data.frame` that contains the demographic information as specified in the `trial_demo` argument of the `clean_sciat()` function. 

Each of the `List` elements can be stored in a separate object. 

```{r}
sciat1 <- sciat_data[[1]] # extract first SC-IAT data
sciat2 <- sciat_data[[2]] # extract second SC-IAT data
demo_data <- sciat_data[[3]] # extract demographic information

head(sciat1)
head(demo_data)
```

### Compute the SC-IAT D

Once the SC-IAT(s) have been prepared and cleaned through the `clean_sciat()` function, it is possible to compute the *D* by means of the `Dsciat()` function. 

This function works by specifying the data set on which the *D* has to be computed (`data`) and the labels identifying the two critical conditions (`mappingA` and `mappingB`). If a rtw was included in the SC-IAT administration, the labels for the responses exceeding the threshold must be specified via the `non_response` argument. 

```{r}
# Compute the D for the first SC-IAT
 d_sciat1 <- Dsciat(sciat1,
                  mappingA = "test.sc_dark.Darkbad",
                  mappingB = "test.sc_dark.Darkgood",
                  non_response = "alert")

# dataframe containing the SC-IAT D of the of the first SC-IAT
str(d_sciat1) 
 
# Compute D for the second SC-IAT
 d_sciat2 <- Dsciat(sciat2,
                  mappingA = "test.sc_milk.Milkbad",
                  mappingB = "test.sc_milk.Milkgood",
                  non_response = "alert")
 
 # dataframe containing the SC-IAT D of the of the second SC-IAT
 head(d_sciat2)

```

The `Dsciat()` function results in a `data.frame` with class `dsciat` containing a number of rows equal to the number of participants, their *D*, and a bunch of useful information on their performance (see the documentation for the `Dsciat()` function for further information on the resulting data frame). Functions `dscr_d()`, `d_plot()`, `d_distr()`, and `sc_grp()` require the object resulting from the `Dsciat()` function to work. 

### Descriptive statistics

The descriptive statistics of the *D* can be easily obtained by means of the `descript_d()` function: 

```{r}
descript_d(d_sciat1)
```

The descriptive statistics of the response times in the two critical conditions are reported as well. 

By specifying `latex = TRUE`, the `dscr_d()` function prints the results in Latex: 

```{r}
descript_d(d_sciat2, latex = TRUE)
```

### Plotting the results

`implicitMeasures` provides the users with different functions for plotting the results, either at the individual or at the sample level. Furthermore, it gives the chance to plot the *D* for multiple SC-IATs. 

#### Individual level 

`d_plot()` plots the *D* for each participant. Participants can be arranged in different ways by means of the `order_sbj` argument. In the default representation, they follow the order they had in the original data frame. By setting the `order_sbj` argument equal to either `D-increasing` or `D-decreasing`, they are arranged by increasing or decreasing *D*, respectively. The values labels of the x-axis can be removed by setting `x_values = FALSE` (suggested in case of big sample size). The `include_stats` (default = `FALSE`) argument plots the descriptive statics of the *D* (mean +/- 2sd). 

```{r fig.align='center', fig.width=6}
 d_plot(d_sciat1)
# change respondents order, remove x-values
d_plot(d_sciat1, order_sbj = "D-increasing", 
       x_values = FALSE)
# change respondents order, remove x-values, add descriptive statistics
d_plot(d_sciat1, order_sbj = "D-increasing", 
       x_values = FALSE, include_stats = TRUE)
```

Points color can be changed as well by using the argument `col_point`. 

#### Sample level

`d_distr()` plots the distribution of the *D*s. It provides different options for choosing the most appropriate distribution by means of the `graph` argument, that can be set equal to `histogram` (histogram plot, default), `density` (density plot), or `violin` (violin plot). Descriptive statistics can be added as well (`include_stats = TRUE`). The filling color for the histogram and the density plot can be modified by means of the `col_fill` argument (e.g., `col_fill = "seagreen"`). The `col_point` argument can be used to change the color of the points in the violin plot (e.g., `col_fill = "royalblue"`).


```{r fig.align='center', fig.width=6}
d_distr(d_sciat1)
# change the number of bins
d_distr(d_sciat1, n_bin = 120)
# change graph and add descriptive statistics
d_distr(d_sciat1, graph = "density", include_stats = TRUE)
#change graph and add descriptive statistics
d_distr(d_sciat1, graph = "violin", include_stats = TRUE)
```

#### Multiple SC-IATs

`multi_dsciat()` plots the distributions of the *D*s from two different SC-IATs. It provides different options for choosing the most appropriate graphical representation by using the `graph` argument and setting it equal to `density` (density plot, default), `violin` (violin plot), or `point` (point plot). The labels differentiating the two SC-IATs can be set by means of the `label_sc1` and `label_sc2` arguments (default are `SC-IAT 1` and `SC-IAT 2`). The values labels of the x-axis can be removed by setting `x-values = FALSE` (suggested in case of big sample size). The default palette of the graph is `dark`. Palette can be changed via `gcolors` argument (available palettes are `greens`, `blues`, and `pinks`).  `dens_mean` add the mean to the density plot (default is `TRUE`). 

```{r fig.align='center', fig.width=6}
multi_dsciat(d_sciat1, d_sciat2)
# remove mean adn change color
multi_dsciat(d_sciat1, d_sciat2, dens_mean = FALSE, 
       gcolors = "pinks")
# change graph
multi_dsciat(d_sciat1, d_sciat2, graph = "violin")
# change graph and labels for the two SC-IATs, remove x-values
multi_dsciat(d_sciat1, d_sciat2, graph = "point", 
       x_values = FALSE, gcolors = "greens",
       label_sc1 = "Dark SC-IAT", 
       label_sc2 = "Milk SC-IAT")
```


